\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-07-18 paper.qmd
Update the choice of bins

```{r}
# Code to draw illustration for notation
start_pt <- all_centroids_df |> 
  filter(hexID == 1)
d_rect <- tibble(x1min = 0, 
                 x1max = 1,
                 x2min = 0,
                 x2max = r2)
#c_start <- tibble(s1 = 0.03, s2 = 0.05)
c_start <- tibble(s1 = 0.03, s2 = -0.05)

@@

              aes(xmin = x1min - c_start$s1,# - c_start$s1, 
                  xmax = x1max - c_start$s1,# - c_start$s1, 
                  ymin = x2min - c_start$s2,# - c_start$s2, 
                  ymax = x2max - c_start$s2),# - c_start$s2), 

@@

    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
                                y=x2min - c_start$s2)) + 
    geom_point(data=d_rect, aes(x=x1max - c_start$s1, 
                                y=x2min - c_start$s2)) + 
    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
                                y=x2max - c_start$s2)) + 
    annotate("text", x=d_rect$x1min - c_start$s1, 
                     y=d_rect$x2min - c_start$s2,

@@

    annotate("text", x=d_rect$x1max - c_start$s1, 
                     y=d_rect$x2min - c_start$s2,

@@

    annotate("text", x=d_rect$x1min - c_start$s1, 
                     y=d_rect$x2max - c_start$s2,

@@

      x = x1min  - c_start$s1, # 0 - 0.03, 

@@

      xend = x1max - c_start$s1, #1 - 0.03, 

@@

      y = x2min - c_start$s2, #0 - 0.05, 

@@

      yend = x2max - c_start$s2), #r2 - 0.05),

@@

<!--add three choice of bins 7, 10, 14-->

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve1 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 7, 
  r2 = r2)

@@

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve2 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.07)

@@

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve3 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 14, 
  r2 = r2,
  q = 0.05)

@@

```{r}
#| echo: false
#| label: fig-bins-scurve
#| fig-pos: H
#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 70$ (7, 10), (b) $b = 240$ (12, 20), and (c) $b = 322$ (14, 23). Color indicates standardized counts, dark indicating high count and light indicates low count. At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins. Using the MSE of the model fit in $p-D$ helps decide on a useful choice of number of bins."
#| fig-width: 6
#| fig-height: 2.5

@@

<!--To generate errors for different total number of bins-->
```{r}
## To initialize number of bins along the x-axis
bin1_vec_scurve <- 4:19 #sqrt(NROW(training_data_scurve)/2)

@@

  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2

  scurve_model <- fit_highd_model(
    training_data = training_data_scurve,
    emb_df = umap_scurve_scaled,
    bin1 = xbins,
    r2 = r2,
    q = 0.07, 
    is_bin_centroid = TRUE,
    is_rm_lwd_hex = FALSE,
    col_start_highd = "x"
  )

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 240, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  geom_vline(xintercept = 96, linetype=2, 
             color = "black", linewidth=0.8, alpha = 0.5) +
  geom_vline(xintercept = 375, linetype=2, 
             color = "black", linewidth=0.8, alpha = 0.5) +
  ylab("log(MSE)") +
  xlab("b") +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 5),
        axis.title.y = element_text(size = 5, 
            angle=90),
        title = element_text(size = 72))
```

>>>>

```{r}
# Code to draw illustration for notation
start_pt <- all_centroids_df |> 
  filter(hexID == 1)
d_rect <- tibble(x1min = 0, 
                 x1max = 1,
                 x2min = 0,
                 x2max = r2)

# To move the rectangle to ignore the overlap with the centroids
rect_adj <- tibble(x1 = 0.03, x2 = 0.03)

@@

              aes(xmin = x1min - rect_adj$x1,# - rect_adj$s1, 
                  xmax = x1max - rect_adj$x1,# - rect_adj$s1, 
                  ymin = x2min - rect_adj$x2,# - rect_adj$s2, 
                  ymax = x2max - rect_adj$x2),# - rect_adj$s2), 

@@

    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
                                y=x2min - rect_adj$x2)) + 
    geom_point(data=d_rect, aes(x=x1max - rect_adj$x1, 
                                y=x2min - rect_adj$x2)) + 
    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
                                y=x2max - rect_adj$x2)) + 
    annotate("text", x=d_rect$x1min - rect_adj$x1, 
                     y=d_rect$x2min - rect_adj$x2,

@@

    annotate("text", x=d_rect$x1max - rect_adj$x1, 
                     y=d_rect$x2min - rect_adj$x2,

@@

    annotate("text", x=d_rect$x1min - rect_adj$x1, 
                     y=d_rect$x2max - rect_adj$x2,

@@

      x = x1min  - rect_adj$x1, # 0 - 0.03, 

@@

      xend = x1max - rect_adj$x1, #1 - 0.03, 

@@

      y = x2min - rect_adj$x2, #0 - 0.05, 

@@

      yend = x2max - rect_adj$x2), #r2 - 0.05),

@@

<!--add three choice of bins 8, 11, 14-->

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve1 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 8, 
  r2 = r2)

@@

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve2 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 11, 
  r2 = r2,
  q = 0.075)

@@

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve3 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 14, 
  r2 = r2,
  q = 0.065)

@@

```{r}
#| echo: false
#| label: fig-bins-scurve
#| fig-pos: H
#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 112 \\text{ } (8, \\text{ }14)  \\text{ }(q = 0.1)$, (b) $b = 220 \\text{ } (11, \\text{ }20) \\text{ }(q = 0.075)$, and (c) $b = 322 \\text{ } (14, \\text{ }26) \\text{ }(q = 0.065)$. Color indicates standardized counts, dark indicating high count and light indicates low count. At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins. Using the MSE of the model fit in $p-D$ helps decide on a useful choice of number of bins."
#| fig-width: 6
#| fig-height: 2.5

@@

<!--To generate errors for different total number of bins-->
```{r}
## To initialize number of bins along the x-axis
bin1_vec_scurve <- 2:19 #sqrt(NROW(training_data_scurve)/2)

@@

  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2

  scurve_model <- fit_highd_model(
    training_data = training_data_scurve,
    emb_df = umap_scurve_scaled,
    bin1 = xbins,
    r2 = r2,
    q = 0.1, 
    is_bin_centroid = TRUE,
    is_rm_lwd_hex = FALSE,
    col_start_highd = "x"
  )

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 220, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  geom_vline(xintercept = 112, linetype=2, 
             color = "black", linewidth=0.8, alpha = 0.5) +
  geom_vline(xintercept = 350, linetype=2, 
             color = "black", linewidth=0.8, alpha = 0.5) +
  ylab("log(MSE)") +
  xlab("b") +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 5),
        axis.title.y = element_text(size = 5, 
            angle=90),
        title = element_text(size = 72))
```

### 2024-07-18 paper.tex
Update the choice of bins

\caption{\label{fig-bins-scurve}Hexbin density plots of UMAP layout of
the S-curve data, using three different bin inputs: (a) \(b = 70\) (7,
10), (b) \(b = 240\) (12, 20), and (c) \(b = 322\) (14, 23). Color
indicates standardized counts, dark indicating high count and light
indicates low count. At the smallest bin size the data segregates into
two separate groups, suggesting this is too many bins. Using the MSE of
the model fit in \(p-D\) helps decide on a useful choice of number of
bins.}

>>>>

\caption{\label{fig-bins-scurve}Hexbin density plots of UMAP layout of
the S-curve data, using three different bin inputs: (a)
\(b = 112 \text{ } (8, \text{ }14)  \text{ }(q = 0.1)\), (b)
\(b = 220 \text{ } (11, \text{ }20) \text{ }(q = 0.075)\), and (c)
\(b = 322 \text{ } (14, \text{ }26) \text{ }(q = 0.065)\). Color
indicates standardized counts, dark indicating high count and light
indicates low count. At the smallest bin size the data segregates into
two separate groups, suggesting this is too many bins. Using the MSE of
the model fit in \(p-D\) helps decide on a useful choice of number of
bins.}
